% !TeX root = ../msc_thesis_jayd.tex
\chapter*{Contributions}
\phantomsection\addcontentsline{toc}{section}{Contributions}
This short section humbly presents my contributions to the 
literature in the form of both conference proceedings and 
peer-reviewed journal articles. 

\section*{Peer-Reviewed Journal Articles}
\renewcommand*{\thefootnote}{\fnsymbol{footnote}}
\stepcounter{footnote}

\begin{enumerate}
 \item[\cite{GAP2014}] G. Painchaud-April\footnote{\label{note1}Both authors contributed equally to this work.}, 
	\textbf{J. Dumont}\textsuperscript{\ref{note1}}, D. Gagnon and L. J. Dubé,\\ 
	``Scattering method for resonances in dielectric microcavities,'' \textit{in preparation}.
  \item[\cite{GAG2014a}] D. Gagnon, \textbf{J. Dumont}, J.-L. Déziel and L. J. Dubé, \\
	``Ab initio investigation of lasing thesholds in photonic molecules,'' \textit{submitted for publication.}
	\href{http://arxiv.og/abs/1401.7305}{\texttt{arXiv:1401.7305}}.
  \item[\cite{GAG2013c}]D. Gagnon, \textbf{J. Dumont} and L. J. Dubé, \\
	``Multiobjective optimization in integrated photonics design,''\\
	Opt. Lett \textbf{38}, 2181--2184 (2013). 
	\href{http://dx.doi.org/10.1364/OL.38002181}{\texttt{doi:10.1364/OL.38002181}}.
  \item[\cite{GAG2012}]D. Gagnon, \textbf{J. Dumont} and L. J. Dubé, \\
	``Beam shaping using genetically optimized two-dimensional photonic crystals,''\\
	J. Opt. Soc. Am. A \textbf{29}, 2673--2678 (2012). 
	\href{http://dx.doi.org/10.1364/JOSAA.29.002673}{\texttt{doi:10.1364/JOSAA.29.002673}}
\end{enumerate}

\clearpage
\section*{Conference Proceedings}
\renewcommand*{\thefootnote}{\roman{footnote}}
\setcounter{footnote}{0}

\begin{enumerate}
 \item[\cite{GAG2014b}] D. Gagnon, \textbf{J. Dumont}, J.-L. Déziel and L. J. Dubé\footnote{Speaker at this invited talk.}, \\
	``Adding SALT to coupled microcavities: the making of active photonic molecule lasers,''\\
	in IEEE Proceedings of the 16th International Conference on Transparent Optical Networks (ICTON 2014).
	\href{http://arxiv.org/abs/1405.0168}{\texttt{arXiv:1405.0168}}
 \item[\cite{GAP2013b}] G. Painchaud-April, \textbf{J. Dumont}\footnote{Speaker at this invited talk.}, D. Gagnon and L. J. Dubé, \\
	``$S$ and $Q$ matrices reloaded: applications to open, inhomogeneous, and complex cavities,''\\
	in IEEE Proceedings of the 15th International Conference on Transparent Optical Networks (ICTON 2013).
	\href{http://dx.doi.org/10.1109/ICTON.2013.6602811}{doi:ICTON.2013.6602811}.
 \item[\cite{GAG2013a}] D. Gagnon\footnote{Speaker at this regular talk.}, \textbf{J. Dumont} and L. J. Dubé, \\
	``Coherent beam shaping using two-dimensional photonic crystals,''\\
	in IEEE Proceedings of the 15th International Conference on Transparent Optical Networks (ICTON 2013).\\
	\href{http://dx.doi.org/10.1109/ICTON.2013.6602751}{\texttt{doi:ICTON.2013.6602751}}.
 \item[\cite{GAG2013b}] D. Gagnon\footnote{Poster presentation}, \textbf{J. Dumont} and L. J. Dubé, \\
	``Optimization in optical systems revisited: beyond genetic algorithms,''\\
	in Bull. Am. Phys. Soc., \textbf{58}(6), 2013. \href{http://meetings.aps.org/link/BAPS.2013.DAMOP.D1.14}{BAPS.2013.DAMOP.D1.14}
\end{enumerate}


