% !TeX root = ../msc_thesis_jayd.tex
\chapter*{Foreword}
\phantomsection\addcontentsline{toc}{section}{Foreword}
\renewcommand*{\thefootnote}{\fnsymbol{footnote}}
\stepcounter{footnote}
This essay, even though the front page claims that it is mine alone, 
is actually the product of a number of people without who it
could not have existed\footnote{Of course, any issue that comes up
is fully mine!}. First and foremost, I must acknowledge
the constant support and trust of Pr. Louis J. Dubé, the head
of our little research group we lovingly call Dynamica. Without his calm
disposition, sharp mind and seemingly infinite knowledge, I could not have done it.
In the two years I have been part of the group, he has helped
me grow both personally and intellectually. I also thank my 
co-supervisor, Younès Messaddeq, for making this essay possible.
Lastly, I wish to thank Pr. Reinhold Blümel; his infectious 
enthusiasm and deep understanding of physics have spawned 
many great ideas. 


I also want to thank a former member of Dynamica, 
Guillaume Painchaud-April, whose footsteps I have
tried to follow. Despite the fact that he had already left
the group when I joined, the conversations we had have helped
make it through this journey. Denis Gagnon deserves a special mention
for teaching me all the tricks in the trade of being a student
researcher. His quick wit, his determination to see things through
(however trivial they might be) 
and  his friendly demeanour have made him both an excellent tutor 
and a fast friend. It goes without saying that all the other
members of Dynamica have been instrumental in the making 
of this essay, and I thank them each individually: Antoine 
Allard, Jean-Luc Déziel, Laurent Hébert-Dufresne, Edward
Laurence and Jean-Gabriel Young, thank you. 

These acknowledgments would not be complete without
mentioning my family. I want to thank my parents 
for their understanding, their support and for 
encouraging my never-ending curiosity. I also 
want to thank my parents-in-law, whose support 
have enabled the completion of many a crazy project, 
including this work. Most of all, however, I want to 
thank my wife, Jessica Cournoyer, whose support and 
understanding have been heavily solicited in the last few years. 
Your love and support have been invaluable in the writing of these words. 

\begin{flushright}
Joey Dumont\\
July 10\textsuperscript{th}, 2014
\end{flushright}